Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-05 method.tex
Switched order of time slices in [ , ) intervals
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@937 ccdd05d5-56c1-461a-bf98-85f6876523b4

For concreteness and simplicity, we will consider an inspiral waveform in the
quadrupole approximation, for which the time-frequency relation is
%
\begin{equation} \label{eq:fgw}
%
f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
\right]^{3/8}.
%
\end{equation}
%
Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G
M_\odot / c^3 \approx 5 \umu\mathrm{s}$) and $t$ is the time relative to the
coalescence of the binary~\cite{findchirppaper, kidder1992}.  Usually the
template is truncated at some prescribed time $t^0$, or equivalently frequency
$f_\textrm{hi}$.  This is often chosen to correspond to the \ISCO. An inspiral
signal will enter the detection band at a low frequency, $f = f_\mathrm{low}$,
corresponding to a time $t_\mathrm{low}$.  The template is assumed to be zero
outside the interval $[t_\mathrm{low}, t^0)$ and is said to have  a duration of
$t^0 - t_\mathrm{low}$. It is critically sampled at a rate of $2
f_\mathrm{hi}$.

The monotonic time-frequency relationship of equation~\eqref{eq:fgw} allows us
to choose time-slice boundaries that require substantially less bandwidth at
early times in the inspiral.  Our goal is to reduce the filtering cost of a
large fraction of the waveform by computing part of the convolution at a lower
sample rate.  Specifically we consider here time slice boundaries with the
smallest power-of-two sample rate that sub-critically samples the time sliced
template.  The time slices consist of the $S$ intervals $\left(t^S,
t^{S-1}\right],\, \dots,\, \left(t^2, t^1\right],\, \left(t^1, t^0\right]$
sampled at frequencies $f^\mathrm{S-1},\, \dots,\, f^1,\, f^0$ where $f^0
\geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.  The time
sliced templates may be downsampled without aliasing, so we define them as
%
\begin{equation}
\label{eq:time-sliced-templates}
h_{i}^{s}[k] \equiv
	\cases{
		h_{i}\!\left[k\frac{f}{f^s}\right] & if $t^s \leqslant k/f^s < t^{s+1}$ \\
		0 & otherwise.
	}
\end{equation}
%
\begin{comment}
An example time-slice design satisfying these constraints for a $1.4 - 1.4 \,
M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
%
\begin{table}[h!]
\begin{minipage}[c]{0.52\textwidth}
\centering
\vspace{0.8cm}
\includegraphics{time_slices.pdf}
\end{minipage}
\begin{minipage}[c]{0.3\textwidth}
\centering
\input{time_slices.tex}
\end{minipage}
\caption{\label{table:time_slices} Example of nearly critically sampled,
power-of-two time slices for a $1.4 - 1.4 \, M_{\odot}$ template extending from
$f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{hi} = f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$
with a time-frequency structure given by ($\ref{eq:fgw})$. $f^s$ is the sample
rate of the time slice, $(t^{s+1}, t^s]$ are the boundaries in seconds
preceeding coalescence and \slicessamps\ is the number of samples in the
$s^{\mathrm{th}}$ filter.}
\end{table}
\end{comment}

>>>>

For concreteness and simplicity, we will consider an inspiral waveform in the
quadrupole approximation, for which the time-frequency relation is
%
\begin{equation} \label{eq:fgw}
%
f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
\right]^{3/8}.
%
\end{equation}
%
Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G
M_\odot / c^3 \approx 5 \umu\mathrm{s}$) and $t$ is the time relative to the
coalescence of the binary~\cite{findchirppaper, kidder1992}.  Usually the
template is truncated at some prescribed time $t^0$, or equivalently frequency
$f_\mathrm{hi}$.  This is often chosen to correspond to the \ISCO. An inspiral
signal will enter the detection band at a low frequency, $f = f_\mathrm{low}$,
corresponding to a time $t_\mathrm{low}$.  The template is assumed to be zero
outside the interval $[t^0, t_\mathrm{low})$ and is said to have  a duration of
$t_\mathrm{low} - t^0$. It is critically sampled at a rate of $2
f_\mathrm{hi}$.

The monotonic time-frequency relationship of equation~\eqref{eq:fgw} allows us
to choose time-slice boundaries that require substantially less bandwidth at
early times in the inspiral.  Our goal is to reduce the filtering cost of a
large fraction of the waveform by computing part of the convolution at a lower
sample rate.  Specifically we consider here time slice boundaries with the
smallest power-of-two sample rate that sub-critically samples the time sliced
template.  The time slices consist of the $S$ intervals
$\left[t^0, t^1\right),\, \left[t^1, t^2\right),\, \dots,\, \left[t^{S-1}, t^S\right)$,
sampled at frequencies $f^0,\, f^1,\, \dots,\, f^\mathrm{S-1}$ where $f^0
\geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.  The time
sliced templates may be downsampled without aliasing, so we define them as
%
\begin{equation}
\label{eq:time-sliced-templates}
h_{i}^{s}[k] \equiv
	\cases{
		h_{i}\!\left[k\frac{f}{f^s}\right] & if $t^s \leqslant k/f^s < t^{s+1}$ \\
		0 & otherwise.
	}
\end{equation}
%
\begin{comment}
An example time-slice design satisfying these constraints for a $1.4 - 1.4 \,
M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
%
\begin{table}[h!]
\begin{minipage}[c]{0.52\textwidth}
\centering
\vspace{0.8cm}
\includegraphics{time_slices.pdf}
\end{minipage}
\begin{minipage}[c]{0.3\textwidth}
\centering
\input{time_slices.tex}
\end{minipage}
\caption{\label{table:time_slices} Example of nearly critically sampled,
power-of-two time slices for a $1.4 - 1.4 \, M_{\odot}$ template extending from
$f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{hi} = f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$
with a time-frequency structure given by ($\ref{eq:fgw})$. $f^s$ is the sample
rate of the time slice, $(t^{s+1}, t^s]$ are the boundaries in seconds
preceeding coalescence and \slicessamps\ is the number of samples in the
$s^{\mathrm{th}}$ filter.}
\end{table}
\end{comment}
